{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-10 thesis/related.tex
rephrase some things in related to make citations not begin on a fresh line, as dave wishes

The problem of relaxed memory nondeterminism alone has inspired the creation of several new tools in the past few years.
Relacy \cite{relacy}, a header-only C++ model checking library for checking synchronization primitives,
was the first to broach this field,
although \revisionminor{it} requires custom annotations for non-atomic memory accesses
and
%(according to later citing papers) % cdschecker
does not fully model all possible relaxed memory behaviours.
%designed for verifying synchronization primitives,
CDSChecker \cite{cdschecker} extends DPOR with a {\em reads-from} relation
to capture most of the C++11 memory model's new behaviours.
Nidhugg \cite{nidhugg} is a checker for TSO and PSO which instruments LLVM abstract assembly,
% i don't actually know what the difference is #overlyhonestmethods
although does not yet support the C++11 memory model.
%% actually it doesn't -- they say they take the Source DPOR worse version. im confus?
% and extends the Optimal DPOR algorithm (discussed below) to include store buffer nondeterminism.
rInspect \cite{tsopso}
%models TSO and PSO differently, using shadow threads, and
offers further heuristic state space reduction using buffer bounding (described below).
RCMC \cite{rcmc} models a ``repaired'' version of the C++11 memory model known as RC11 \cite{rc11},
and professes to achieve the best state space reduction to date.
These tools each use various heuristics to account for spin-wait loops,
ranging from delay bounding \cite{bpor} to a rigid rewrite rule,
and provide only limited support so far for read-modify-write atomics
(at best, supporting them by introducing % oops
some redundant exploration).
%
No relaxed-memory MC has yet proposed a satisfactory model for the ``thin-air'' problem \cite{sully-thesis},
which can cause state space cycles in a way not yet well-understood and remains future work.
%and notably includes relaxed memory nondeterminism in its concurrency model.
They also identify all data races
(under the C++ definition rather than \cref{sec:quicksand-soundness}'s; see \cref{sec:background-hb})
as bugs immediately, rather than checking them for benign or buggy outcomes.
All the tools in this paragraph are notably open-source -- an encouraging recent trend in the field.

@@

Several other recent advances extend DPOR to new concurrency models,
beyond the shared-memory-threading model outlined in \cref{sec:landslide-dpor}.
TransDPOR \cite{transdpor} provides extra domain-specific reduction for message-passing actor programs
by exploiting the fact that the dependency relation is transitive in the absence of shared state.
The $R^4$ algorithm \cite{r4} (corresponding to the R4 checker mentioned above)
extends DPOR to event-driven programs by separating the notion of enabled events from that of multiple threads.
TaxDC \cite{taxdc}, a taxonomy study of distributed systems concurrency bugs,
showed that for completeness distributed model checkers must incorporate many forms of nondeterminism,
including message reordering, timeouts, network disconnections, and crashes and reboots, in addition to local threads.
DPOR for TSO and PSO \cite{tsopso}
extends the concurrency model
using {\em shadow threads}, which interleave with traditional threads to represent store buffer nondeterminism,
which can expose bugs not even possible in the strong consistency model
such as discussed in \cref{sec:tm-warpzone-relaxed}.
It also introduced a heuristic {\em buffer bounding} technique, analogous to ICB,
to mitigate the corresponding increase in state space size.
The same year, Nidhugg \cite{nidhugg} proposed a DPOR extension to account for TSO and PSO
using {\em chronological traces}.
MCR was recently extended to support relaxed memory models likewise \cite{mcr-tsopso}.
Just this year, RCMC \cite{rcmc} proposed to replace the interleaving model entirely with {\em execution graphs},
which precisely model the executions legal under the RC11 memory model,
offering further reduction still.
Somewhat analogously for HTM, this work's \Cref{chap:tm}
extended DPOR's concurrency model to include failure injection,
% not "new" -- abstraction reduction is the same as always
and proposed three reduction strategies, one sound and two heuristic,
%to make some progress up the exponential mountain.
to keep state spaces manageable.

@@

Since these foundational algorithms,
many more recent works have contributed to make data-race analysis more precise, more performant,
and/or more domain-specific.
The Causally-Precedes relation \cite{predictive-dr} is a refinement of Limited HB which avoids the most common cases of false positives,
including \cref{sec:quicksand-realloc}'s reallocation false positives.
It could strike a middle ground in the bug-finding/verification tradeoff
between Pure and Limited HB (\cref{sec:quicksand-eval})
that would be a welcome enhancement in Quicksand.
IFRit \cite{ifrit}
improves the performance of Pure HB using an interference analysis,
which could allow future work to avoid tracing every memory access in a simulator such as Bochs \cite{bochs} or Simics \cite{simics}.
%
DroidRacer \cite{droidracer} and CAFA \cite{cafa} find data races in Android applications,
using domain-specific heuristics (orthogonal to Quicksand's method) to reduce false positives.
DataCollider \cite{datacollider} finds data races in kernel code
by using hardware breakpoints and random sampling to achieve high performance.
% TODO: read some more recent confs for data race shit since thesprop
% probably like weak memory c.c

@@

%% this 2nd paper... not freely available.......
%while Kolikant \cite{learning-concurrency} investigates how students form cognitive patterns about concurrent programming that could either aid or stunt their reasoning.
%Both of these studies could help optimize Landslide's bug reports for clarity and student enlightenment.
Willgrind \cite{willgrind} is a tool recently developed at Virginia Tech
that targets a fork-join parallelism project
and checks for memory errors (using the Valgrind \cite{valgrind} framework)
as well as deadlocks, assertion failures, and data races, similarly to Landslide,
although unlike Landslide, its thread interleaving coverage is as yet limited to stress testing.
Its GUI-based debugging output is perhaps more friendly than Landslide's HTML preemption traces,
and its user survey found that students appreciated detailed debugging info especially for deadlocks
(future work for Landslide),
but also that students had little patience for even a 5-minute stress test
when no assurance against false negatives could be provided.
This suggests motivating students with Landslide's verification guarantee,
% TODO fix this explanation its too roundabout
although it is tricky to avoid accientally encouraging % oops
them to limit possible interleavings by just using one global lock for everything,
which is counter to 15-410's educational goals.

@@

Transactional memory (TM), first introduced in 1993 \cite{transactional-memory},
% TODO expand this to talk about some stm implementations?
has received renewed attention in recent years with the announcement of Intel's Haswell architecture \cite{htm-haswell},
which supports hardware transactions (HTM) using new x86 instructions.
Since then, many studies have evaluated the increased performance it offers over traditional locking and/or STM
\cite{htm-experience,htm-performance,tm-benchmark-cmu}.
%
HTM's performance comes at an increased cost in complexity to the programmer,
who must avoid system calls or transaction nesting, respect the CPU cache capacity,
and consider retry loops for spurious failure.
SI-TM \cite{si-tm} introduces techniques for reducing HTM's abort rates for performance's sake,
but without eliminating them altogether, any full verification must still consider them possible anywhere.
For programmers who wish to avoid such concerns,
the simpler STM programming model remains relevant.
One recent work \cite{hybrid-htm-stm} enhances STM transactions to nest with HTM ones,
while another \cite{stm-relaxed-memory} adds support for relaxed memory models.
Meanwhile, two recent papers \cite{relaxed-transactions-pldi,relaxed-transactions-popl}
have proposed formal models of HTM's execution semantics under relaxed memory likewise.
Such extensions come with the challenge of even more complicated behavioural semantics
for MCs to accurately model and verify in future work.

@@

%An article from relatively early in the timeline of TSX
\revisionminor{The paper which originally defined weak and strong atomicity~\cite{htm-subtleties}
also}
warns of several false equivalence pitfalls when converting conventionally-locking code to use transactions,
although these pitfalls depend on multiple existing locks used locally and disjointly,
so this does not invalidate the equivalence proved in \cref{sec:tm-design-locks}.
Rather, Landslide could be used to ensure that freshly-converted transactional code avoids the warned-of pitfalls.
{\em Learning from Mistakes} \cite{learning-from-mistakes},
a survey of the characteristics of many types of concurrency bugs,
found that TM could potentially fix some, but not all, of the studied bugs,
while in other cases it must be combined with other concurrency primitives to be fully correct.
A subsequent paper \cite{applying-tm-bugs}
found a majority of bugs in their study to be easily fixable with hand-written transactions,
while others remained out of scope due to blocking {\tt cond\_wait()} operations and the like;
more recently, the tool BugTM \cite{bugtm} aims to deploy such repairs in production code fully automatically.
However, these studies all optimize for empirical correctness at best,
as well as maintaining good performance,
which motivates the use of tools like Landslide to ensure these rewrites
are actually correct,
% oops
rather than merely shrinking the necessary preemption window required to expose them.

@@

While C's extremely rudimentary type system allows the compiler to statically check programs
for properties such as not accidentally dereferencing raw integer values as if they were pointers,
more advanced programming languages may make guarantees about concurrent execution.
%
\revision{Language-level concurrency abstractions were first introduced by the Communicating Sequential Processes (CSP)
model~\cite{communicating-sequential-processes},
which defines a concurrent program as a set of isolated processes
allowed to communicate only via blocking input/output commands.
Erlang \cite{erlang}, an early concurrent functional language,
introduced the actor model for concurrency based on the CSP model,
in which threads likewise} share no state and must communicate only by message-passing.
%One might say this statically guarantees freedom from data races;
%from another point of view, data races do not even exist as a concept in this programming model.
While this statically guarantees the absence of data races,
programs may still execute nondeterministically,
so concurrency bugs, especially deadlocks, are not ruled out.
Concuerror \cite{concuerror}, discussed above, is a MC tool for Erlang programs.
%
Haskell \cite{haskell} offers a more sophisticated interface to concurrency:
threads may reference the same objects and even update shared references
using monads that encapsulate mutation,
but at the (garbage-collected) execution level all data is immutable once created,
which preserves type soundness and data-race freedom.
The aforementioned D\'{e}j\`{a} Fu \cite{dejafu} checks concurrent Haskell programs.
%
Rust %\cite{rust-language}
presents a type system with more explicit memory management,
in-place mutation,
and mutable references
to appear familiar and approachable to those already versed in C++.
It proposes a borrow-check analysis to ensure memory and type safety
despite mutable references \cite[\S{}4.2,\S{}10.3]{rust-book},
and a trait system to ensure no shared state between threads by default \cite[\S{}10.2,\S{}16.4]{rust-book}.
Its concurrency libraries then offer interfaces which relax this restriction,
allowing threads even to simultaneously reference shared mutable state,
using the type system to enforce sound locking discipline across such accesses,
again preserving type soundness and data-race freedom \cite[\S{}16.2]{rust-book}.%
\footnote{The author themself contributed the original design for this latter feature \cite{rust-rwarc}.}
I know of no existing model checker for Rust as of yet.
%
The Relaxed Memory Calculus \cite{sully-thesis}
proposes to extend C++ with annotations for weak memory atomics,
which allows for static formal analysis of memory access reorderings.
Although not ruling out data races,
this approach is an important step towards compilers
which can statically reason about program execution under more advanced concurrency models.
%
Finally, LVish \cite{lvish}
features a type system that enforces deterministic behaviour by construction,
using shared state called LVars which allow writes only in ways that update order is not observable.
This renders thread interleavings entirely irrelevant,
obviating any need for runtime verification,
but at the cost of a more restrictive programming model.

@@

Analogous to stateless MC,
which seeks good coverage of possible thread execution paths under schedule nondeterminism,
another popular testing approach is symbolic execution \cite{symbolic},
which seeks good coverage of possible flow control paths under input nondeterminism.
Symbolic executors
abstract a program's variables and use constraint solvers such as Z3~\cite{z3}
to work backwards and synthesize combinations of test inputs which can lead to a failure.
%
KLEE \cite{klee}, one well-known and open-source implementation,
offers over 90\% code coverage on average across many tests,
often outdoing that achieved by programmers' own hand-written tests.
%
Later, Contessa \cite{contessa} extended symbolic execution to include concurrency nondeterminism as well,
by using a DPOR-like analysis on individual execution traces
then including reordering possibilities in its SMT constraints.
%rather than using explicit scheduling as MC does.
This simultaneously exercises both input and schedule nondeterminism,
%and achieves good bug-finding performance in practice,
but does not provide the same verification guarantees as repeated DPOR iterations with explicit scheduling.
Exploring both kinds of state space at once thoroughly enough to provide strong verification
is undoubtedly subject to further state space explosion,
and remains future work.
%
Symbiosis \cite{symbiosis} starts from the known root cause of an existing failure
and uses symbolic execution to synthesize a schedule to reproduce it,
then further searches for a non-failing schedule and compares them to produce
a minimum sequence of events necessary for the failure.
This approach skips the initial verification step entirely,
but greatly reduces the diagnosis effort required of the user,
which was a common complaint about Landslide's preemption traces.

>>>>

The problem of relaxed memory nondeterminism alone has inspired the creation of several new tools in the past few years.
Relacy \cite{relacy}, a header-only C++ model checking library for synchronization primitives,
was the first to broach this field,
although \revisionminor{it} requires custom annotations for non-atomic memory accesses
and
%(according to later citing papers) % cdschecker
does not fully model all memory reorderingss.
%designed for verifying synchronization primitives,
CDSChecker \cite{cdschecker} extends DPOR with a {\em reads-from} relation
to capture most of the C++11 memory model's new behaviours.
Nidhugg \cite{nidhugg} is a checker for TSO and PSO which instruments LLVM abstract assembly,
% i don't actually know what the difference is #overlyhonestmethods
although does not yet support the C++11 memory model.
%% actually it doesn't -- they say they take the Source DPOR worse version. im confus?
% and extends the Optimal DPOR algorithm (discussed below) to include store buffer nondeterminism.
rInspect \cite{tsopso}
%models TSO and PSO differently, using shadow threads, and
offers further heuristic state space reduction using buffer bounding (described below).
RCMC \cite{rcmc} models a ``repaired'' version of the C++11 memory model known as RC11 \cite{rc11},
and professes to achieve the best state space reduction to date.
These tools each use various heuristics to account for spin-wait loops,
ranging from delay bounding \cite{bpor} to a rigid rewrite rule,
and provide only limited support so far for read-modify-write atomics
(at best, supporting them by introducing % oops
some redundant exploration).
%
No relaxed-memory MC has yet proposed a satisfactory model for the ``thin-air'' problem \cite{sully-thesis},
which can cause state space cycles in a way not yet well-understood and remains future work.
%and notably includes relaxed memory nondeterminism in its concurrency model.
They also identify all data races
(under the C++ definition rather than \cref{sec:quicksand-soundness}'s; see \cref{sec:background-hb})
as bugs immediately, rather than checking them for benign or buggy outcomes.
All the tools in this paragraph are notably open-source -- an encouraging recent trend in the field.

@@

Several other recent advances extend DPOR to new concurrency models,
beyond the shared-memory-threading model outlined in \cref{sec:landslide-dpor}.
TransDPOR \cite{transdpor} provides additional domain-specific reduction for message-passing actor programs
by exploiting the fact that the dependency relation is transitive in the absence of shared state.
The $R^4$ algorithm \cite{r4}, used by the R4 checker mentioned above,
extends DPOR to event-driven programs by separating the notion of enabled events from that of multiple threads.
TaxDC \cite{taxdc}, a taxonomy study of distributed systems concurrency bugs,
showed that for completeness distributed model checkers must incorporate many forms of nondeterminism,
including message reordering, timeouts, network disconnections, and crashes and reboots, in addition to local threads.
DPOR for TSO and PSO \cite{tsopso}
extends the concurrency model
using {\em shadow threads}, which interleave with traditional threads to represent store buffer nondeterminism,
which can expose bugs not even possible in the strong consistency model
such as discussed in \cref{sec:tm-warpzone-relaxed}.
It also introduced a heuristic {\em buffer bounding} technique, analogous to ICB,
to mitigate the corresponding increase in state space size.
The same year, Nidhugg \cite{nidhugg} proposed a DPOR extension to account for TSO and PSO
using {\em chronological traces}.
MCR was recently extended to support relaxed memory models likewise \cite{mcr-tsopso}.
Just this year, RCMC \cite{rcmc} proposed to replace the interleaving model entirely with {\em execution graphs},
which precisely model the executions legal under the RC11 memory model,
offering further reduction still.
Somewhat analogously for HTM, this work's \Cref{chap:tm}
extended DPOR's concurrency model to include failure injection,
% not "new" -- abstraction reduction is the same as always
and proposed three reduction strategies, one sound and two heuristic,
%to make some progress up the exponential mountain.
to keep state spaces manageable.

@@

Since these foundational algorithms,
many more recent works have contributed to make data-race analysis more precise, more performant,
and/or more domain-specific.
The Causally-Precedes relation \cite{predictive-dr} is a refinement of Limited HB which avoids the most common cases of false positives,
including \cref{sec:quicksand-realloc}'s reallocation false positives.
It could strike a middle ground in the bug-finding/verification tradeoff
between Pure and Limited HB (\cref{sec:quicksand-eval})
that would be a welcome enhancement in Quicksand.
IFRit \cite{ifrit}
improves the performance of Pure HB using an interference analysis,
which could allow future work to avoid tracing every memory access in a simulator such as Bochs \cite{bochs} or Simics \cite{simics}.
%
DroidRacer \cite{droidracer} and CAFA \cite{cafa} find data races in Android applications,
using domain-specific heuristics (orthogonal to Quicksand's method) to reduce false positives.
DataCollider \cite{datacollider} finds data races in kernel code
by using hardware breakpoints and random sampling to achieve high performance.

@@

%% this 2nd paper... not freely available.......
%while Kolikant \cite{learning-concurrency} investigates how students form cognitive patterns about concurrent programming that could either aid or stunt their reasoning.
%Both of these studies could help optimize Landslide's bug reports for clarity and student enlightenment.
Willgrind \cite{willgrind} is a tool recently developed at Virginia Tech
that targets a fork-join parallelism project
and checks for memory errors (using the Valgrind \cite{valgrind} framework)
as well as deadlocks, assertion failures, and data races, similarly to Landslide,
although unlike Landslide, its thread interleaving coverage is as yet limited to stress testing.
Its GUI-based debugging output is perhaps more friendly than Landslide's HTML preemption traces,
and its user survey found that students appreciated detailed debugging info especially for deadlocks
(future work for Landslide),
but also that students had little patience for even a 5-minute stress test
when no assurance against false negatives could be provided.
This suggests motivating students with Landslide's verification guarantee,
although it is tricky to avoid accientally encouraging % oops
them to limit possible interleavings by just using one global lock for everything,
which is counter to 15-410's educational goals.

@@

Transactional memory (TM), first introduced in 1993 \cite{transactional-memory},
has received renewed attention in recent years
%with the announcement of
since the launch of Intel's Haswell architecture \cite{htm-haswell},
which supports hardware transactions (HTM) using new x86 instructions.
Since then, many studies have evaluated the increased performance it offers over traditional locking and/or STM
\cite{htm-experience,htm-performance,tm-benchmark-cmu}.
%
HTM's performance comes at an increased cost in complexity to the programmer,
who must avoid system calls or transaction nesting, respect the CPU cache capacity,
and consider retry loops for spurious failure.
SI-TM \cite{si-tm} introduces techniques for reducing HTM's abort rates for performance's sake,
but without eliminating them altogether, any full verification must still consider them possible anywhere.
For programmers who wish to avoid such concerns,
the simpler STM programming model remains relevant.
One recent work \cite{hybrid-htm-stm} enhances STM transactions to nest with HTM ones,
while another \cite{stm-relaxed-memory} adds support for relaxed memory models.
Meanwhile, two recent papers \cite{relaxed-transactions-pldi,relaxed-transactions-popl}
have proposed formal models of HTM's execution semantics under relaxed memory likewise.
Such extensions come with the challenge of even more complicated behavioural semantics
for MCs to accurately model and verify in future work.

@@

%An article from relatively early in the timeline of TSX
\revisionminor{The paper which originally defined weak and strong atomicity~\cite{htm-subtleties}
also}
warns of several false equivalence pitfalls when converting conventionally-locking code to use transactions,
although these pitfalls depend on multiple existing locks used locally and disjointly,
so this does not invalidate the equivalence proved in \cref{sec:tm-design-locks}.
Rather, Landslide could be used to ensure that freshly-converted transactional code avoids the warned-of pitfalls.
{\em Learning from Mistakes} \cite{learning-from-mistakes},
a survey of the characteristics of many types of concurrency bugs,
found that TM could potentially fix some, but not all, of the studied bugs,
while in other cases it must be combined with other concurrency primitives to be fully correct.
A subsequent paper \cite{applying-tm-bugs}
found a majority of bugs to be easily fixable with hand-written transactions,
while others remained out of scope due to blocking {\tt cond\_wait()} operations and the like;
more recently, the tool BugTM~\cite{bugtm} aims to deploy such repairs in production code fully automatically.
However, these studies all optimize for empirical correctness at best,
as well as maintaining good performance,
which motivates the use of tools like Landslide to ensure these rewrites
are actually correct,
% oops
rather than merely shrinking the necessary preemption window required to expose them.

@@

While C's extremely rudimentary type system allows the compiler to statically check programs
for properties such as not accidentally dereferencing raw integer values as if they were pointers,
more advanced programming languages may make guarantees about concurrent execution.
%
\revision{Language-level concurrency abstractions were first introduced by the Communicating Sequential Processes (CSP)
model~\cite{communicating-sequential-processes},
which defines a concurrent program as a set of isolated processes
allowed to communicate only via blocking input/output commands.
Erlang \cite{erlang}, an early concurrent functional language,
introduced the actor model for concurrency based on the CSP model,
in which threads likewise} share no state and must communicate only by message-passing.
%One might say this statically guarantees freedom from data races;
%from another point of view, data races do not even exist as a concept in this programming model.
While this statically guarantees the absence of data races,
programs may still execute nondeterministically,
so concurrency bugs, especially deadlocks, are not ruled out.
Concuerror \cite{concuerror}, discussed above, is a MC tool for Erlang programs.
%
Haskell \cite{haskell} offers a more sophisticated interface to concurrency:
threads may reference the same objects and even update shared references
using monads that encapsulate mutation,
but at the (garbage-collected) execution level all data is immutable once created,
which preserves type soundness and data-race freedom.
The aforementioned D\'{e}j\`{a} Fu \cite{dejafu} checks concurrent Haskell programs.
%
Rust's
%\cite{rust-language}
type system supports more explicit memory management,
in-place mutation,
and mutable references
to appear familiar and approachable to those already versed in C++.
It proposes a borrow-check analysis to ensure memory and type safety
despite mutable references \cite[\S{}4.2,\S{}10.3]{rust-book},
and a trait system to ensure no shared state between threads by default \cite[\S{}10.2,\S{}16.4]{rust-book}.
Its concurrency libraries then offer interfaces which relax this restriction,
allowing threads even to simultaneously reference shared mutable state,
using the type system to enforce sound locking discipline across such accesses,
again preserving type soundness and data-race freedom \cite[\S{}16.2]{rust-book}.%
\footnote{The author themself contributed the original design for this latter feature \cite{rust-rwarc}.}
I know of no existing model checker for Rust as of yet.
%
The Relaxed Memory Calculus \cite{sully-thesis}
proposes to extend C++ with annotations for weak memory atomics,
which allows for static formal analysis of memory access reorderings.
Although not ruling out data races,
this approach is an important step towards compilers
which can statically reason about program execution under more advanced concurrency models.
%
Finally, LVish \cite{lvish}
features a type system that enforces deterministic behaviour by construction,
using shared state called LVars which allow writes only in ways that update order is not observable.
This renders thread interleavings entirely irrelevant,
obviating any need for runtime verification,
but at the cost of a more restrictive programming model.

@@

Analogous to stateless MC,
which tests many possible thread execution paths under schedule nondeterminism,
another popular testing approach is symbolic execution \cite{symbolic},
which tests many possible flow control paths under input nondeterminism.
Symbolic executors
abstract a program's variables and use constraint solvers such as Z3~\cite{z3}
to work backwards and synthesize combinations of test inputs which can lead to a failure.
%
KLEE \cite{klee}, one well-known and open-source implementation,
offers over 90\% code coverage on average across many tests,
often outdoing that achieved by programmers' own hand-written tests.
%
Later, Contessa \cite{contessa} extended symbolic execution to include concurrency nondeterminism as well,
by using a DPOR-like analysis on individual execution traces
then including reordering possibilities in its SMT constraints.
%rather than using explicit scheduling as MC does.
This simultaneously exercises both input and schedule nondeterminism,
%and achieves good bug-finding performance in practice,
but does not provide the same verification guarantees as repeated DPOR iterations with explicit scheduling.
Exploring both kinds of state space at once thoroughly enough to provide strong verification
is undoubtedly subject to further state space explosion,
and remains future work.
%
Symbiosis \cite{symbiosis} starts from the known root cause of an existing failure
and uses symbolic execution to synthesize a schedule to reproduce it,
then further searches for a non-failing schedule and compares them to produce
a minimum sequence of events necessary for the failure.
This approach skips the initial verification step entirely,
but greatly reduces the diagnosis effort required of the user,
which was a common complaint about Landslide's preemption traces.
